Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2020-09-20 manuscript/ERL_global_C_review.Rmd
worked on manuscript (mostly results)

Forest ecosystems globally play a critical role in regulating atmospheric carbon dioxide (CO~2~), and thereby will shape the course of climate change (IPCC1.5). 
Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**). 
While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades. 
The resulting CO~2~ sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). 
Moreover, forests contain substantial carbon (C) stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C [biomass and soils; @bonan_forests_2008]. 
Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**). 
The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. 
Regrowth forests in particular will play an important role [@pugh_role_2019], as these represent a large (~#%) and growing proportion of Earth's forests.
Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; Cavaleri et al 2015].

@@

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature). 
Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd). 
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006]. 
Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
Records for which stand age was unknown (#% of total database) were excluded from the analysis. 
*These groupings were defined based on ForC v.2.0 [@anderson-teixeira_forc_2018], for which they covered 91% of the primary variable records for forests of known age, or 86% of total records. The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.*

We drew upon records for *#* annual flux and *#* C stock variables (Table 1). For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Table 1, [this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))). 
Although ForC contains information that may be used to standardize or control for methodological differences [*e.g.*, area sampled, min stem diameter sampled, allometric equations applied; @anderson-teixeira_forc_2018], for this analysis we included all relevant data in ForC. 
Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm. All records were measured directly or derived from field measurements (as opposed to modeled).

@@

Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses. 
In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). 
Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX A*. 
Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age). 
Specifically, we removed all plots flagged as managed in ForC-simplified (managed field). 
This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.  
Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”). 
We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods. 
We also removed all plots for which no stand history information had been retrieved.

Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages following an approach similar to that of @anderson-teixeira_carbon_2016. 
For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot, weighting flux measurements according to the length of measurement periods (*i.e.*, end.date - start.date). 
Values were then averaged across plots clustered within 25 km of one another (geographic.area field of SITES table, sensu Anderson-Teixeira et al 2018), weighting by area.sampled (MEASUREMENTS table) or plot.area (PLOTS table) if available for all records. 
This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC. 
Finally, for figures 6 and 7, the original values were analyzed via a linear mixed effects model ('lmer' function in 'lme4' R package) with biome as fixed effect and plot nested within geographic.area as random effects on the intercept. 
When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. There were enough data to run this analysis for all focal variables but **biomass_ag_understory, BNPP_root.turnover_fine, deadwood_down, total.ecosystem_2.**

@@

## Review Results/ Synthesis
```{r eval=FALSE, echo=FALSE}
# The Review Results/Synthesis section presents the findings of the analyses.
```
### Data Coverage

Of the *#* records in ForC v.#, *#* met the criteria for inclusion in this study (Fig. 1). 
*(Give some stats on coverage by age and biome)*

@@

Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5, and statistics for each biome type are also summarized at: [GitHub URL]. 
Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests. 
For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent. 
That is, component fluxes and stocks summed to within 1 std of more inclusive fluxes in all but one instance (in temperate conifer forests, $aboveground woody biomass + foliage biomass > aboveground biomass + 1std$; Fig. 5). **update this: https://github.com/forc-db/ERL-review/issues/16 **

@@

(**check paragraph with latest data**) The largest C fluxes—including $GPP$, $NPP$, $ANPP$, $BNPP$, $R_{soil_het}$, $R_{soil}$, $R_{eco}$, — were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests  ([ForC_variable_averages_per_Biome](https://github.com/forc-db/ForC/blob/master/numbers_and_facts/ForC_variable_averages_per_Biome.csv)) (Fig. 6). 
The same held true for some of the subsidiary fluxes: ANPP_foliage, ANPP_woody, [**OTHERS?**]. Other subsidiary fluxes—including NPP_woody, ANPP_repro, ANPP_stem, ANPP_branch, woody.mortality, BNPP_coarse, BNPP_fine,  [**OTHERS?**]—deviated from this pattern and/or lacked data for some biomes. 
Net ecosystem production ($NEP$) did not follow this pattern, with no significant differences across biomes but the largest average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance ($NEP$).

```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$. Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively). Lines...  .", out.width='100%', fig.pos='H'}

@@

There were less distinct trends in C stocks across biomes (Fig. 7). 
Biome means for live aboveground and total biomass followed the same general trend as the major fluxes, with biomass of tropical= temperate broadleaf $\ge$ temperate conifer> boreal forests (Fig. 7a). 
However, the relative differences in these means were much smaller than for the major C fluxes, and there was significant variation within biomes. 
Maximum aboveground biomass values followed a very different trend than the means: temperate broadleaf> temperate conifer>boreal>tropical. 
There were some statistically significant biome differences in less frequently sampled C stocks (*e.g.*, woody biomass, foliage biomass, deadwood), but given high within-biome variability in C stocks and relatively low sample sizes, these were likely attributable to sampling biases and methodological differences than to true differences across biomes.

```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.", out.width='100%', fig.pos='H'}

@@

(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 8-11. In general, ForC contained roughly comparable information on C fluxes in regrowth vs mature forests, with the notable exception of tropical forests, for which there were no fluxes that had been measured in $\ge$ 7 different geographic areas. C stocks were better documented for regrowth stands, with a total of # records from # distinct geographic areas. Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Fig. 5-10; detailed below). Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs. 6-7).

>>>>

Forest ecosystems will play a critical role in shaping the course of climate change (IPCC1.5) through their influence on atmospheric carbon dioxide (CO~2~). 
Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**). 
While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades. 
The resulting CO~2~ sink averaged 3.0 ± 0.8 Gt C yr^{-1} from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). 
Moreover, forests contain substantial carbon (C) stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C [biomass and soils; @bonan_forests_2008]. 
Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**). 
The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. 
Regrowth forests in particular will play an important role [@pugh_role_2019], as these represent a large (~#%) and growing proportion of Earth's forests.
Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; Cavaleri et al 2015].

@@

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature). 
Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd). 
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006]. 
Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
Records for which stand age was unknown (#% of total database) were excluded from the analysis. 
These groupings covered *#%* of the primary variable records for forests of known age, or *#%* of total records. 
The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with *#* records total.

We selected 23 annual flux and 11 C stock variables for inclusion in the analysis, although two flux variables ($R_{het-ag}$ and $R_{het}$) were included for conceptual completeness but had no records in ForC  (Table 1). 
For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Tables 1, S#(variable mapping)). 
Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm. All records were measured directly or derived from field measurements (as opposed to modeled).

@@

Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses. 
In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). 
Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX S1*. 
Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age). 
Specifically, we removed all plots flagged as managed in ForC-simplified (managed field). 
This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.  
Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”). 
We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods. 
We also removed all plots for which no stand history information had been retrieved.

Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages. **(this par & following are confusing/ incomplete--needs work)**
For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot. 
Values were then averaged across plots clustered within 25 km of one another [*sensu* @anderson-teixeira_forc_2018], weighting by area sampled if available for all records. 
This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC. 
Finally, for figures 6 and 7, the original values were analyzed via a linear mixed effects model ('lmer' function in 'lme4' R package) with biome as fixed effect and plot nested within geographic.area as random effects on the intercept. 
When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. There were enough data to run this analysis for all focal variables but $DW_{down}$.

@@

## Review Results/ Synthesis
```{r eval=FALSE, echo=FALSE}
# The Review Results/Synthesis section presents the findings of the analyses.
```

*NOTE TO COAUTHORS: THIS SECTION HAS A TON OF FIGURES, BUT VERY LITTLE TEXT. IS THERE ANYTHING THAT SHOULD BE ADDED?*

### Data Coverage

Of the *#* records in ForC v.#, *#* met the criteria for inclusion in this study (Fig. 1). 
These records were distributed across *#* plots in *#* distinct geographic areas. 
Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests. 
For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests.

@@

Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5 (and available in tabular format in the ForC release accompanying this publication).

For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent. 
That is, component fluxes and stocks summed to within one standard deviation of more inclusive fluxes in all but two instances, both for living C stocks in temperate conifer forests (Fig. 5). **(check this with final results: https://github.com/forc-db/ForC/blob/master/numbers_and_facts/C_cycle_closure.csv) ** 
Within this biome, $B_{ag}<B_{ag-wood}+B_{foliage}$ and $B_{root}<B_{root-coarse}+B_{root-fine}$ because $B_{ag-wood}$ and $B_{root-coarse}$ were very high, with strongly disproportionate numbers of records from the high-biomass forests of the US Pacific Northwest (Figs. S18, S21).

@@

**(The following covers only the variables with age trends figures; there maybe a few more without much age data)**
With only # exceptions, C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs. 6, S1-S15).
In this majority of cases, differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both. 
Fluxes tended to be greater in temperate broadleaf than conifer forests, but the difference was never significant.
This pattern held for the following variables: $GPP$, $NPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{foliage}$, $ANPP_{litterfall}$, $BNPP$, $BNPP_fine$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$.  
The most notable exception to this pattern was $NEP$, with no significant differences across biomes but with the largest average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests (Figs. 5,S1).
Another exception was for $BNPP_{root-coarse}$, where all records came from high-biomass forests in the US Pacific Northwest, resulting in high values for the temperate conifer biome and no significant differences across biomes (Fig. S10).  
Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, with the important exception in the overall C balance ($NEP$).

```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$. Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively). Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome. Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes. Individual figures for each flux with sufficient data given in the Supplement (Figs. S1-S15).", out.width='100%', fig.pos='H'}

@@

There were fewer distinct trends in C stocks across biomes (Figs. 7, S16-S26). 
Specifically, there were significant differences for only *#* of *#* variables: *LIST*. 
In all of these cases, temperate conifer forests were in the highest significance grouping, and boreal forests in the lowest. 
Tropical and temperate broadleaf forests fell in between--most commonly being significantly different from temperate conifer but not boreal forests.
... 
*(These differences are driven more by geographic distribution of sampling than by true differences.)*

```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood. Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively). Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome. Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes. Individual figures for each stock with sufficient data given in the Supplement (Figs. S16-S26).", out.width='100%', fig.pos='H'}

@@

(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 8-11.  Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Table 1, Figs. 5-10, S1-S26; detailed below). Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs. 6-7).
